A Trio of 40 to 50 Million Year Old Mini-Neptunes from Kepler, TESS, and Gaia

### 2022-04-07 papers/ms.tex
add lithium appendix

>>>>

\section{Spectroscopic Youth Indicators}
\label{app:spectra}

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f5a.pdf}
			\includegraphics[width=0.49\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.3cm}
	\caption{
    {\bf Lithium 6708\AA\ equivalent widths for Kepler-1643, KOI-7368,
    KOI-7913 AB, and Kepler-1627 compared to those of young open
    clusters and field stars. }
    The statistical uncertainties on the newly measured equivalent
    widths are shown, or else are smaller than the markers.
    The field stars are stars with Kepler planets from \citet{berger_identifying_2018}.
    The ``40-50 Myr'' reference stars (left) are from
		IC\,2602 \citep{randich_gaiaeso_2018} and Tuc-Hor
		\citep{kraus_stellar_2014}.
    The ``112 Myr'' stars (right) are from the Pleiades
    \citep{soderblom_evolution_1993,jones_evolution_1996,bouvier_pleiades_lirot_2018},
    where we adopt the LDB age from \citet{dahm_2015}.
    \label{fig:lithium}
	}
\end{figure*}

\subsection{Lithium}

Figure~\ref{fig:lithium} compares the measured lithium equivalent
widths of the objects of interest against a few reference populations.
We selected the reference populations from the literature only for
cases in which upper limits were explicitly reported.  KOI-7368 and
KOI-7913~A have secure lithium detections, and the KOI-7913~B
detection is marginal (${\rm EW}_{\rm Li} = 42^{+12}_{-19}$).  For all
three stars, as well as for Kepler-1627~A, the observed lithium
equivalent width is consistent with the stellar effective temperatures
and $\approx40$\,Myr ages.

Kepler-1643, in RSG-5, is conspicuously below the 40-50$\,$Myr sequence
in the left-panel of Figure~\ref{fig:lithium}, while still being above
the field stars (${\rm EW}_{\rm Li} = 130^{+6}_{-5}$).
The right panel shows the comparison against the Pleiades, where
Kepler-1643 is more consistent with the observed dispersion in lithium.

One explanation for this could be that Kepler-1643 is a field
interloper; another could be that RSG-5 is older than 50\,Myr.  We do
not favor the former, given that {\it i)} the star was selected based
on its spatio-kinematic proximity to other RSG-5 members, nearly all
of which appear consistent with being $\approx$50\,Myr old
(Figure~\ref{fig:age} upper left), and {\it ii)} the same RSG-5
members display a gyrochronal sequence consistent with being younger
than the Pleaides (Figure~\ref{fig:age} lower left).
We do not have an explanation, but in the following paragraph
speculate on one possible line of thinking that might help resolve the
discrepancy.

Analyzing Figure~\ref{fig:lithium}, it is remarkable that in just 50
million years, stars between $4500$\,K and $5200$\,K go from having
such a tight lithium sequence to one with a dispersion
$\approx10\times$ greater.  The existence of such a dispersion in
Pleiades-age K-dwarfs has been known for longer than some of the
present authors have been alive; it has also been known that the stars
with the most lithium are also the most rapidly rotating
\citep{butler_pleiades_1987,soderblom_evolution_1993}.  More recent
analyses of this correlation have been reviewed by
\citet{bouvier_lithium-rotation_2020}.  The conclusion of that work
was that the origin of the rotation-lithium correlation likely lies
within pre-main-sequence stellar physics.  If so, one would expect the
IC\,2062 and Tuc-Hor K-dwarfs to show a larger intrinsic lithium
dispersion, which they do not.  It seems therefore that the exact
onset time of the K-dwarf lithium dispersion might be worth
clarifying.  \citet{messina_rotation-lithium_2016} for instance do not
obviously see it in the $\beta$~Pic moving group ($25\pm3$\,Myr).
Given that RSG-5 seems somewhat isochronally older than the
$\delta$~Lyr cluster (and therefore IC\,2602
\citealt{bouma_kep1627_2022}), it might be a worthy cluster for such
an analysis.
